---
title: "STAT 652 Assignment 1"
author: "Dhruv Patel"
date: "13/10/2021"
output: pdf_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

Lecture 5 Application         
We will now do variable selection with these five variables     
1. Use all-subsets regression.      
```{r} 
rm(list=ls(all=TRUE))
data = na.omit(airquality[,1:4])

data$TWcp = data$Temp*data$Wind
data$TWrat = data$Temp/data$Wind

filter_data = model.matrix(Ozone ~.,data= data)
head(filter_data)

library(leaps)
allsub <- regsubsets(x=filter_data, 
                     y=data$Ozone, intercept = F)

info.subsets = summary(allsub)
seq.subsets = info.subsets$which
vars.seq.subsets.raw = apply(seq.subsets, 1, function(W){
  vars.list = names(W)[W]
  output = paste0(vars.list, collapse = ", ")})
```
(a) Report the variables in the best model of each size.      
```{r}
print(vars.seq.subsets.raw)
```


(b) Compute BIC on each of these models and report the BIC values for the
models.          
Answer:       
```{r}
print(info.subsets$bic)
```

(c) Identify the best model. What variables are in it?           
Answer: The best model is (Intercept),Solar.R,Temp,TWrat since its has minimum BIC value i.e. -207.1195            
```{r}
print(min(info.subsets$bic))
```

2. Use the hybrid stepwise algorithm that is the default in the step() function. Report
the model that it chooses as “best.”     
Answer: The best model according to stepwise algorithm is TWrat + Temp + Solar.R.          
```{r}
data$TWcp = data$Temp*data$Wind
data$TWrat = data$Temp/data$Wind
head(data)
rows = nrow(data)

initial <- lm(data=data, formula=Ozone~ 1)
final <- lm(data=data, formula=Ozone~Solar.R+Wind+Temp+TWcp+TWrat)
step <- step(object=initial, scope=list(upper=final), k = log(rows))
summary(step)
```

3. Use 10-fold CV to estimate the MSPE for the stepwise model selection process. That
is,       
(a) Set the seed to 2928893 before running the sample.int() function.       
(b) Create 10 folds     
(c) Run step() on each training set      
(d) Find the best model, and compute the prediction error on it       
(e) Report the separate MSPEs from each fold, MSP Ev, v = 1, . . . , 10 and
the MSPE for the full data.     
Answers:       
```{r}
set.seed(2928893)
rows = nrow(data)
V=10
folds = floor((sample.int(rows)-1)*V/rows) + 1 
mat_CV_L5 = matrix(NA, nrow=V, ncol=1)

for(v in 1:V){
  
  initial <- lm(data=data[folds != v,], formula=Ozone~ 1)
  final <- lm(data=data[folds != v,], formula=Ozone~Solar.R+Wind+Temp+TWcp+TWrat)
  rows = nrow(data[folds != v,])
  step <- step(object=initial, scope=list(upper=final), k = log(rows))
  
  pred = predict(step,newdata=data[folds==v,])
  summary(pred)
  mat_CV_L5[v,1] = mean((data[folds==v,"Ozone"] - pred)^2)
}

lm_best = lm(Ozone ~ TWrat + Temp + Solar.R, data = data)
summary(lm_best)
plot(lm_best)

colnames(mat_CV_L5) = c('Each fold MPSE')
mat_CV_L5
mean(mat_CV_L5) 
```


