---
title: "STAT 652 Assignment 1"
author: "Dhruv Patel"
date: "13/10/2021"
output: pdf_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

Lecture 4 Application B       
1. Once zone and make have been converted to factors, run the linear regression with per
as the response and the other six variables as explanatory.      
(a) Create a summary of the lm object.     
i. Although you fit a model with 6 variables, how many parameters are
estimated?    
Answer:     
```{r}
ins = read.csv('/Users/dhruv/Downloads/Insurance-1.csv',header=TRUE)
ins$zone = as.factor(ins$zone)
ins$make=as.factor(ins$make)

ins = ins[ins$claims>0,]
dim(ins)

ins_per_lm = lm(per ~ ., data = ins)
summary(ins_per_lm)
```

```{r}
length(ins_per_lm$coefficients)
```
ii. What is the intercept of the regression when make and zone are both
at their first level, 1?     
Answer: To find intercept for first level everything else will be zero. So finally we will get intercept = 11.86


iii. What is the intercept of the regression when make and zone are both
at their last levels, 9 and 7, respectively?      
Answer: To find intercept at make9 and zone7 will be (make7) =  (Intercept) + (Intercept at make9) + (Intercept at zone 7) = 11.86 - 2.862 - 1.716 = 7.282     
