---
title: "STAT 652 Assignment 1"
author: "Dhruv Patel"
date: "13/10/2021"
output: pdf_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

Lecture 4: Applications A       
1. Compute a summary on TWcp and TWrat. Report the minimum, maximum, and
mean for each variable.    
Answer:
```{r}
data <-  read.csv("/Users/dhruv/Downloads/midterm.csv",header=TRUE, sep=",", na.strings=" ")

data = na.omit(data)
filter_data = (data[,1:4])
head(filter_data)

filter_data$TWcp = filter_data$Temp*filter_data$Wind
filter_data$TWrat = filter_data$Temp/filter_data$Wind
```

Min, Max and Mean values for TWcp are:      
```{r}
min(filter_data$TWcp)
max(filter_data$TWcp)
mean(filter_data$TWcp)
```

Min, Max and Mean values for TWrat are:
```{r}
min(filter_data$TWrat)
max(filter_data$TWrat)
mean(filter_data$TWrat)
```

2. Create two new models: Temp + Wind + TWcp and Temp + Wind + TWrat. Fit these
two models in lm().      


(a) Report the t-test results for the two new variables.      
Answer:      
TWrat summary:      
```{r}
lm_twrat = lm(Ozone ~ Temp + Wind + TWrat, data = filter_data)
summary(lm_twrat)
```

TWcp summary:     
```{r}
lm_twcp = lm(Ozone ~ Temp + Wind + TWcp, data = filter_data)
summary(lm_twcp)
```

(b) Based on the test results, which variable seems to be the most useful, or are
neither particularly helpful? (1 sentence)      
Answer:     
t values for TWcp and TWrat are -3.987 and 4.005 respectively.Since both values are below significance level(0.5). Both are important.      

(c) From the model with the cross-product term, compute and report the slope
of the Temp effect when Wind is at its minimum value. Repeat for the
maximum value of Wind. (You can do this by hand from the output if you
want.)
```{r}
min(filter_data$Wind)
max(filter_data$Wind)
```
                  
                                    
                                                      
                                                                                          
                  
                  
                                    
                                                      
                                                                                          
                  
                                    
                                    
                                                      
                                                                                          
                  
                                    
                                    
                                                      
                                                                                          
                  
3. Fit each model on the training data and report the MSPEs from the validation
data.      
(a) Which model wins this competition?     
Answer:       
```{r}
set.seed(2928893)
rows = nrow(filter_data)
train_split = 0.75
reorder_col = sample.int(n=rows, size=rows, replace=FALSE)
set = ifelse(test = ((train_split*rows) > reorder_col), yes=1, no=2)

train_data = filter_data[set==1,]
test_data = filter_data[set==2,]

fit.TWcp = lm(Ozone ~ Temp + Wind + TWcp, data = train_data)
fit.TWrat = lm(Ozone ~ Temp + Wind + TWrat, data = train_data)

pred.TWcp = predict(fit.TWcp, newdata=test_data)
pred.TWrat = predict(fit.TWrat,newdata=test_data)

MSPE.TWcp = mean((test_data$Ozone - pred.TWcp)^2)
MSPE.TWrat = mean((test_data$Ozone - pred.TWrat)^2)

MSPE.TWcp
MSPE.TWrat
```
So,when set.seed(2928893) is fixed. We get MSPE for TWcp as 286.4392 and TWrat as 290.9852. Which shows TWcp wins the competition.       


4. Add these models the five you compared in the previous exercise, and rerun the CV
20 times.         
(a) Make boxplots of the RMSPE, and narrow focus if necessary to see
best models better.         
Answer: V = 7 corresponding to different models and R = 20 number of times it runs.        
```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

data$TWcp = data$Temp * data$Wind
data$TWrat = data$Temp / data$Wind

V=7 
R=20 

mat_CV = matrix(NA, nrow=V*R, ncol=7)
colnames(mat_CV) = c("Solar.R", "Wind", "Temp","TWcp","TWrat","All","Custom")

for (i in 1:R){ 
  folds = floor((sample.int(rows)-1)*V/rows) + 1 
  for(j in 1:V){
    
    # Training Model
    fit.Solar.R = lm(Ozone ~ Solar.R, data = data[folds!=j,])
    fit.Wind = lm(Ozone ~ Wind, data = data[folds!=j,])
    fit.Temp = lm(Ozone ~Temp, data = data[folds!=j,])
    fit.TWcp = lm(Ozone ~ Temp + Wind + TWcp, data = data[folds!=j,])
    fit.TWrat = lm(Ozone ~ Temp + Wind + TWrat, data = data[folds!=j,])
    fit.All = lm(Ozone ~ ., data = data[folds!=j,])
    fit.Custom = lm(Ozone ~ .^2 + Solar.R^2 + Wind^2 + Temp^2, data = data[folds!=j,])
    
    # Model Prediction
    pred.Solar.R = predict(fit.Solar.R, newdata = data[folds==j,])
    pred.Wind = predict(fit.Wind, newdata = data[folds==j,])
    pred.Temp = predict(fit.Temp, newdata = data[folds==j,])
    pred.TWcp = predict(fit.TWcp, newdata = data[folds==j,])
    pred.TWrat = predict(fit.TWrat,newdata = data[folds==j,])
    pred.All = predict(fit.All, newdata = data[folds==j,])
    pred.Custom = predict(fit.Custom,newdata = data[folds==j,])
    
    r = j+V*(i-1)
    # Calculating MSPE for each attributes
    mat_CV[r,1] = mean((data[folds==j,"Ozone"] - pred.Solar.R)^2)
    mat_CV[r,2] = mean((data[folds==j,"Ozone"] - pred.Wind)^2)
    mat_CV[r,3] = mean((data[folds==j,"Ozone"] - pred.Temp)^2)
    mat_CV[r,4] = mean((data[folds==j,"Ozone"] - pred.TWcp)^2)
    mat_CV[r,5] = mean((data[folds==j,"Ozone"] - pred.TWrat)^2)
    mat_CV[r,6] = mean((data[folds==j,"Ozone"] - pred.All)^2)
    mat_CV[r,7] = mean((data[folds==j,"Ozone"] - pred.Custom)^2)
  }
}
```

MSPE Cross-Validation Boxplot:       
```{r}
# MSPE Boxplot
boxplot(mat_CV, las=2, ylim=c(0,1300),main="MSPE Cross-Validation")
```


Relative MSPE Cross-Validation Boxplot:        
```{r}
rel_CV = mat_CV/apply(mat_CV, 1, min)
boxplot(rel_CV, las=2,ylim=c(0,10),main="Relative MSPE Cross-Validation")
```
(b) Are any of the new models competitive, or even best? (1 sentence)        
Answer: The model with second-order for three varialbes (Solar.R, Wind and Temp) is best model till now.     